---
title: "ZX2SAMPLE_STARTING_MARKDOWN - (000) Project Name"
author: "RESEARCH ANALYST"
date: "2024-XX-XX"
output:
  html_document:
    theme: null
    css: ../style.css
    include: 
      before_body: ../header.html
      after_body: ../footer.html
---

# Sample code
[Provide a brief description] Simple template to begin new analayis.

## Changelog
- Programmer:  
- Date:  
- Change Description:  
- Template Update [Y|N]:  
- One-Off:  

## Input data

- dfgjnusql-db71p/Sub_SDS.dbo.[table/sp]
- ../../CSV/05 - Main/[filename]

*NOTE:* No files from file-system are read-in.
*NOTE:* No database extracts are executed.
 
## Output data

- ../../CSV/01 - CSIS/csisDB.db [dat_harvest]
- ../../CSV/05 - Main/[filename]

## Background
[Omit if unecessary]

## Checklist
- Update header information
- Check values
- etc...

## Additional information
Provide additional information if necessary

## Required libraries 
[Adjust as needed]
- tidyVerse  
- knitr

```{r setup, echo=FALSE}

# Set some knit options and functions for formatting data.
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(include = TRUE)
knitr::opts_chunk$set(results='asis')

options(knitr.kable.NA = '')

```

# Data processing

## Initialize environment

Loading libraries, standard functions and set environment parameters.

```{r initalize environment}

# Clear out all existing variables & datasets.
rm(list=ls(all=TRUE))

# Load libraries
library(knitr)

# Include the project parameters file - this needs to be updated for all 
#   projects.
source('../../Z00_PROJECT_PARAMETERS.r')
source('../../Z01_RMD_HTML_FORMATTING_FUNCTIONS.r')

# Note project information
cat(str_interp('<p class="rbn">Working on project ${projID} ${projectName} - ${studyear}</p>'))

```

## Load data

```{r load data}

# ##############################################################################
# Load data from intermediate files.
# ##############################################################################

# This for each file opened.
myData <- read.csv('../../CSV/01 - Database Extract/myFile.csv',
                   na = '', 
                   header = TRUE, 
                   strip.white = TRUE)

count <- nrow(myData)
cat(formatSummaryBlock(paste("Opening file: XXXXXX, ", 
                             count, 
                             " records loaded.", sep="")))


```


## Clean and organize

Write notes about items that must be organized or cleaned and why. Also note any automation that was used.

```{r clean and organize}



```

## Estimates

Provide description of how estimates are derived or any deviations from standard expansions.

```{r estimates}


```

## Reports

```{r write tables and figures}
# This is an example of a table, you can remove the col.names option here
#  and use only the data frame column headers, depending..
knitr::kable(myData,
             caption=formatTableHeader("First ten community codes"),
             col.names=c("List",
                         "Column",
                         "Names",
                         "Here"),
             align="l")

```

## Final clean-up

```{r final organization}

```

## Write SQLite

```{r sql lite export}

# This block can be deleted if SQLite is not being used.

```

## Write CSV

```{r write csv files}

# ##############################################################################
# Write out CSV files.
# ##############################################################################

# Sample code for writing out a file.
  fName = str_interp('../../CSV/01 - Database Extract/${dataSetName}.csv')
  cat(formatSummaryBlock(
    paste('Writing file: ', fName, 
          ' ', nrow(myData), ' records to be written', sep='')))

# Standard is to use CSV files for portability.
  rio::export(myData, fName)

```

<p class="h1footer"> End of Check SDS HH0 script. </p>
